On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2023-10-03 Manuscript/manuscript.tex
Changed some language, still need to fix some things

There are two types of data that are explored throughout this paper.  For the
signed-rank tests described in Section~\ref{sec:rank}, the data comes from 
athletes who competed in the 2022 World Championships and another competition 
(2019 World Championships, 2023 World Championships, or 2022 national 
championships). For the GLMM described in Section~\ref{sec:glmm}, data comes
from every world championships from 1999 to 2023. This investigation started 
shortly after the 2022 World Championships, but as the 2023 data has now become 
available, we are pleased to include it in both of our analyses, and it did not
significantly change our results \citep{WAData}. 
\jy{I added a sentence to the end of intro.}

@@

With our small sample size, the asymptotic normal distribution may not be
reliable, so we also use 1000 random permutations to simulate the null 
distribution of the test statistic.  \jy{increase 1000 to a much bigger number,
  say 10 million}
This method is available from the \texttt{clusWilcox.test()} function
with \texttt{method = `ds'} (for \underline{D}atta and \underline{S}atten) and 
\texttt{exact = TRUE} from R package
\texttt{clusrank} \citep{jiang2020wilcoxon}.

@@

Let $Y_{ijk}$ be the reaction time of observation~$k$ in heat~$j$ in year~$i$.
Conditional on a venue effect $v_i$ of year~$i$ and a heat effect
$h_{i/j}$ nested within each year~$i$, $Y_{ijk}$ is modeled by a
gamma distribution. In the classic generalized linear model (GLM) sense,
$Y_{ijk}$ has conditional mean $E(Y_{ijk} | v_i, h_{i/j}) > 0$ and a
dispersion parameter~$\phi > 0$. The conditional mean is characterized by
\[
g\{E(Y_{ijk} | v_i, h_{i/j})\} = \alpha + v_i + h_{i/j},
\]
where $g$ is a known link function,
$v_i$ is normally distributed with mean zero and
variance~$\sigma_v^2$, and $\delta_{i/j}$ 
is normally distributed with mean
zero and variance~$\sigma_h^2$.
Although \citet{lo2015idlink} suggested that the gamma 
family with an identity link works well for reaction time in cognitive
psychological research, the logarithm
link function worked better for our data and additionally ensures the positivity
of the conditional mean. The gamma mixed-effect model can be fit with the 
\texttt{glmer()} function in R package \texttt{lme4} \citep{lme4}.

@@

Table~\ref{tab:Clusrankresults} shows the results generated from both the 
permutation (exact) and asymptotic rank-based tests for the three comparisons, 
as discussed previously in Section~\ref{sec:rank}.    
The national versus international comparison
produced lower p-values than the 2019 versus 2022 comparison, with the
national vs international permutation test producing a p-value of 
$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  
\eds{Prof. Yan, is this possible based on 1000 permutations?}
\jy{That cannot be true.  Redo with a much bigger number of permutations and
  update the table.}
\of{Done, permutation results were for B = 1 million}
This
is a highly significant test result and shows that athletes who competed at the
2022 World Track and Field Championships were on average, significantly faster
at reacting than they were just a few months prior.  For the 2019 versus 2022
same athlete comparison with permutation test, the p-value was 
$2.9\cdot10^{-3}$, which is significant at $\alpha = 0.01$.  The result of
comparing 2022 and 2023 was very similar to the above result, as the p-value
from the permutation test was $3.0\cdot10^{-3}$.  Taken together, these three
tests help show that athletes competing at the 2022 World Track and Field
Championships had faster reaction times than when they had competed previously.

@@

The estimated parameters of the GLMM model from our analysis are shown in 
Table~\ref{tab:Gamma_parameters}.  These statistics reflect the inclusion of 2022
in our analysis and shows the three models we considered. $\sigma_h^2$ is the
standard deviation of the heat effect, $\sigma_v^2$ is the standard deviation of
the venue effect, and $\alpha$ is the intercept.
The GLMM with both venue and heat random
effects was the best model as evidenced by a decrease in the AIC compared to the
other models with only one random effect. 
Thus, results from the GLMM with both venue and heat random effects
will be presented throughout the rest
of the paper.

\begin{table}
  \centering
  \caption{Estimated parameters for the two effect GLMM on various data sets. SE
  is short for standard error. 
	\eds{This table is not showing what you have described here or in the
          main text.}
        \jy{report $ a = 1 / \phi$}
      } 
  \begin{tabular}{c c c c c c}
      \toprule
      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $1 / \phi$ \\
      \midrule
      Venue Effect Only & $-2023.5$ & N/A & 0.054 & $-1.871$ & 0.150\\
      Heat Effect Only & $-2070.7$ & 0.073 & N/A & $-1.890$ & 0.135\\
      Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
      \bottomrule
  \end{tabular}
  \label{tab:Gamma_parameters}
\end{table}

>>>>

There are two types of data that are explored throughout this paper.  For the
signed-rank tests described in Section~\ref{sec:rank}, the data comes from 
athletes who competed in the 2022 World Championships and another competition 
(2019 World Championships, 2023 World Championships, or 2022 national 
championships). For the GLMM described in Section~\ref{sec:glmm}, data comes
from every world championships from 1999 to 2023. This investigation started 
shortly after the 2022 World Championships, but as the 2023 data has now become 
available, we are pleased to include it in both of our analyses, and it did not
significantly change our results \citep{WAData}.

@@

With our small sample size, the asymptotic normal distribution may not be
reliable, so we also use 1000 random permutations to simulate the null 
distribution of the test statistic.
This method is available from the \texttt{clusWilcox.test()} function
with \texttt{method = `ds'} (for \underline{D}atta and \underline{S}atten) and 
\texttt{exact = TRUE} from R package
\texttt{clusrank} \citep{jiang2020wilcoxon}.

@@

Let $Y_{ijk}$ be the reaction time of observation~$k$ in heat~$j$ in year~$i$.
Conditional on a venue effect $v_i$ of year~$i$ and a heat effect
$h_{i/j}$ nested within each year~$i$, $Y_{ijk}$ is modeled by a
gamma distribution. In the classic generalized linear model (GLM) sense,
$Y_{ijk}$ has conditional mean $E(Y_{ijk} | v_i, h_{i/j}) > 0$ and a
dispersion parameter~$\phi > 0$. The conditional mean is characterized by
\[
g\{E(Y_{ijk} | v_i, h_{i/j})\} = \alpha + v_i + h_{i/j},
\]
where $g$ is a known link function,
$v_i$ is normally distributed with mean zero and
variance~$\sigma_v^2$, and $h_{i/j}$ 
is normally distributed with mean
zero and variance~$\sigma_h^2$.
Although \citet{lo2015idlink} suggested that the gamma 
family with an identity link works well for reaction time in cognitive
psychological research, the logarithm
link function worked better for our data and additionally ensures the positivity
of the conditional mean. The gamma mixed-effect model can be fit with the 
\texttt{glmer()} function in R package \texttt{lme4} \citep{lme4}.

@@

Table~\ref{tab:Clusrankresults} shows the results generated from both the 
permutation (exact) and asymptotic rank-based tests for the three comparisons, 
as discussed previously in Section~\ref{sec:rank}. The national versus international
comparison produced lower p-values than the 2019 versus 2022 comparison, with the
national vs international permutation test producing a p-value of $2\cdot10^{-6}$.
This is a highly significant test result and shows that athletes who competed at the
2022 World Track and Field Championships were on average, significantly faster
at reacting than they were just a few months prior.  For the 2019 versus 2022
same athlete comparison with permutation test, the p-value was 
$2.9\cdot10^{-3}$, which is significant at $\alpha = 0.01$.  The result of
comparing 2022 and 2023 was very similar to the above result, as the p-value
from the permutation test was $3.0\cdot10^{-3}$.  Taken together, these three
tests help show that athletes competing at the 2022 World Track and Field
Championships had faster reaction times than when they had competed previously.

@@

The estimated parameters of the GLMM model from our analysis are shown in 
Table~\ref{tab:Gamma_parameters}.  These statistics reflect the inclusion of 2022
in our analysis and shows the three models we considered. $\sigma_h^2$ is the
standard deviation of the heat effect, $\sigma_v^2$ is the standard deviation of
the venue effect, $\alpha$ is the intercept, and $a$ is the shape parameter of
the gamma function.
The GLMM with both venue and heat random
effects was the best model as evidenced by a decrease in the AIC compared to the
other models with only one random effect. 
Thus, results from the GLMM with both venue and heat random effects
will be presented throughout the rest
of the paper.

\begin{table}
  \centering
  \caption{Estimated parameters for the two effect GLMM on various data sets. SE
  is short for standard error. 
	\eds{This table is not showing what you have described here or in the
          main text.}
        \jy{report $ a = 1 / \phi$}
      } 
  \begin{tabular}{c c c c c c}
      \toprule
      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $a$ \\
      \midrule
      Venue Effect Only & $-2023.5$ & N/A & 0.054 & $-1.871$ & 0.150\\
      Heat Effect Only & $-2070.7$ & 0.073 & N/A & $-1.890$ & 0.135\\
      Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
      \bottomrule
  \end{tabular}
  \label{tab:Gamma_parameters}
\end{table}
